---
title: "Writing Your Journal Article in 12 Weeks"
subtitle: A guide to academic publishing success by Wendy Laura Belcher
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

I'm trying to organize my preparing research for publication, as well as explore the timeline capabilities of R. I have prepared the Table of Contents from Wendy Laura Belcher's "Writing Your Journal Article in 12 Weeks" workbook as a .csv file, previously built via Microsoft Excel. Using a data frame format I found while looking for help with Gantt charts in R, the dataframe can be seen here:

```{r}

df<-read.csv("Table of Contents.csv")
data.frame(df)
head(df)
```

Each of the 12 weeks (with an additional intro week and wrap-up week) has several instructional topics and exercises to perform. The Gantt chart would ideally show a timeline with these items clearly visible, preferably displayed by week. The language I was able to uncover and use to a mild degree of success was this:

```{r}
library(DiagrammeR)
m<-mermaid(
  paste0(
    # mermaid "header", each component separated with "\n" (line break)
    "gantt", "\n", 
    "dateFormat  YYYY-MM-DD", "\n", 
    "title Journal Article Gantt Diagram", "\n",
    # unite the first two columns (task & status) and separate them with ":"
    # then, unite the other columns and separate them with ","
    # this will create the required mermaid "body"
    paste(df %>%
            unite(i, task, status, sep = " : ") %>%
            unite(j, i, pos, start, end, sep = ", ") %>%
            .$j, 
          collapse = "\n"
    ), "\n"
  )
)

m$x$config = list(ganttConfig = list(
  axisFormatter = list(list(
    "%b  %Y"
    ,htmlwidgets::JS(
      'function(d){ return d.getDay() == 1 }'
    )
  ))
))
```
and evenutally, after tweaking my task list to not contain colons or semi-colons and rooting through mistakes in the schedule labelling, I was able to force this out:

```{r}
m
```
While all the tasks are there, this version of a Gantt chart is not ideal, in that it is not obvious how to format the chart to be better viewed. A closer look at this format may result in a more usable image.
```{r}

```
