---
title: "Writing Your Journal Article in 12 Weeks"
subtitle: A guide to academic publishing success by Wendy Laura Belcher
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

I'm trying to organize my preparing research for publication, as well as explore the timeline capabilities of R. I have prepared the Table of Contents from Wendy Laura Belcher's "Writing Your Journal Article in 12 Weeks" workbook as a .csv file, previously built via Microsoft Excel. Using a data frame format I found while looking for help with Gantt charts in R, the dataframe can be seen here:

```{r}

df<-read.csv("Table of Contents.csv")
data.frame(df)
```

Each of the 12 weeks (with an additional intro week and wrap-up week) has several instructional topics and exercises to perform. The Gantt chart would ideally show a timeline with these items clearly visible, preferably displayed by week. The language I was able to uncover and use to a mild degree of success was this:

```{r}
library(DiagrammeR)
m<-mermaid(
  paste0(
    # mermaid "header", each component separated with "\n" (line break)
    "gantt", "\n", 
    "dateFormat  YYYY-MM-DD", "\n", 
    "title Journal Article Gantt Diagram", "\n",
    # unite the first two columns (task & status) and separate them with ":"
    # then, unite the other columns and separate them with ","
    # this will create the required mermaid "body"
    paste(df %>%
            unite(i, task, status, sep = " : ") %>%
            unite(j, i, pos, start, end, sep = ", ") %>%
            .$j, 
          collapse = "\n"
    ), "\n"
  )
)

m$x$config = list(ganttConfig = list(
  axisFormatter = list(list(
    "%b  %Y"
    ,htmlwidgets::JS(
      'function(d){ return d.getDay() == 1 }'
    )
  ))
))
```
and evenutally, after tweaking my task list to not contain colons or semi-colons and rooting through mistakes in the schedule labelling, I was able to force this out:

```{r}
m
```
While all the tasks are there, this version of a Gantt chart is not ideal, in that it is not obvious how to format the chart to be better viewed. A closer look at this format may result in a more usable image.

Another timeline package I came across was "timevis". I manipulated the .csv a bit, inputting actual dates instead of the "after" nomenclature for the Gantt coding. The result is something a bit more polished, being able to zoom in on the task much more easily, and having the timeline in a more landscape feel rather than the overwhelming profile given in the Gantt.

```{r}
library(timevis)

data <- data.frame(
  content = df$task,
  start   = df$start,
  end     = df$end
)
```
Resulting in:
```{r}
timevis(data)
```
Some features I would like to add to this timeline is to color code the tasks by topic and status and to have the information for the task be available when moused over.

Another R timeline package is "plotly".

```{r}
library(plotly)

df$Start  <- as.Date(df$start, format = "%m/%d/%Y")
client    <- "Sample Client"
cols      <- RColorBrewer::brewer.pal(length(unique(df$client)), name = "Set3")
df$color  <- factor(df$client, labels = cols)

p <- plot_ly()
for(i in 1:(nrow(df) - 1)){
  p <- add_trace(p,
                 x = c(df$Start[i], df$Start[i]+df$duration[i]), 
                 y = c(i, i), 
                 mode = "lines",
                 line = list(color = df$color[i], width = 20),
                 showlegend = F,
                 hoverinfo = "text",
                 text = paste(as.character(df$Task[i]), "<br>",
                              "Status: ", df$status[i]),
                 evaluate = T
                 )
                 }

p
```
Adding some formatting might make this one easier to look at:
```{r}
p <- layout(p,
            
            # Axis options:
            # 1. Remove gridlines
            # 2. Customize y-axis tick labels and show task names instead of numbers
            
            xaxis = list(showgrid = F, tickfont = list(color = "#e6e6e6")),
            
            yaxis = list(showgrid = F, tickfont = list(color = "#e6e6e6"),
                         tickmode = "array", tickvals = 1:nrow(df), ticktext = unique(df$task),
                         domain = c(0, 0.9))
)
            
            # Annotations
            
annotations = list(
              # Add total duration and total resources used
              # x and y coordinates are based on a domain of [0,1] and not
              # actual x-axis and y-axis values
              
              list(xref = "paper", yref = "paper",
                   x = 0.80, y = 0.1,
                   text = paste0("Total Duration: ", sum(df$Duration), " days<br>",

              # Add client name and title on top
              
              list(xref = "paper", yref = "paper",
                   x = 0.1, y = 1, xanchor = "left",
                   text = paste0("Gantt Chart: ", client),
                   font = list(color = "#f2f2f2", size = 20, family = "Times New Roman"),
                   ax = 0, ay = 0,
                   align = "left")
            )))
            
            plot_bgcolor = "#333333"  # Chart area color
            paper_bgcolor = "#333333"  # Axis area color
 
p
```

